---
title: "Sustainable Growth: Exploring Integrative Multi-Trophic Aquaculture on the West Coast"
subtitle: '<span style="font-size: 30px; text-align: center; display: block;">Prioritizing Potential Finfish Species</span>'
author: "Jaden Orli"
date: last-modified
bibliography: 
 - ../bibs/references.bib
 - ../bibs/packages.bib
nocite: |
  @tidyverse, @tmap, @here, @janitor, @knitr, @stringr, @tibble, @kableExtra, @calecopal, @htmltools
execute: 
  eval: true
format:
  html: 
    theme: minty
    css: custom.css
    toc: true
    toc-title: "Table of Contents:"
editor_options: 
  chunk_output_type: inline
embed-resources: true
---

# I. Background

## a) Motivation

With the U.N. predicting that the global human population will reach
approximately 9.7 billion people by the year 2050, access to basic
resources (particularly food and water) will continue to become
increasing pressing issues [@un_population2024]. The global assessment
of the state of food security and nutrition in 2022 highlights the
severity of this issue at current population levels (\~8.2 billion):

"It is estimated that between 691 and 783 million people in the world
faced hunger in 2022. Considering the midrange (about 735 million), 122
million more people faced hunger in 2022 than in 2019, before the
pandemic [@fao_food_security2023]."

This highlights the increasingly urgent need to address the lack of
global food security. As a result, marine aquaculture has become the
center of research; with many notable characteristics that make it a
sustainable alternative to meet the growing demands for protein sources
[@hall2011; @nash2021]. Gentry et al. mapped potential locations that
would be suitable for marine aquaculture around the world, and they
found that global seafood demand could be met using less than 0.015% of
the global ocean area [@gentry2017]. Additionally, an analysis of
systems that have adopted an integrated multi-trophic aquaculture (IMTA)
model show promise as a sustainable approach to open-ocean marine
aquaculture [@troell2009].

Gentry et al. identified potential locations based on human and
environmental constraints. This included excluding areas; that would
negatively impact the growth of bivalves (low phytoplanktonic food
availability) and finfish (low dissolve oxygen), were too expensive (too
deep \>200m) to anchor farms, were already allocated to other uses, or
were in high-density shipping areas [@gentry2017].

Based on this analysis, we aim to identify locations along the west
coast of the U.S. that would be optimal for the establishment of a
marine aquaculture farm for both oysters and various finfish species
based on two environmental parameters that impact optimal growth:

1.  range of thermal tolerance
2.  habitat range (based on depth)

## b) Oysters

Research has shown that the following conditions are required for
optimal oyster growth:

1.  sea surface temperature: 11-30°C
2.  depth: 0-70 meters below sea level

## c) Potential Finfish Species {#potential_finfish_species}

Potential finfish species were selected based on the current commercial
fishes managed by the NOAA West Coast operation [@noaa_west_coast].
Highly migratory species (such as sharks) were removed from
consideration resulting in a list of 26 [potential finfish species](#load_species). We then used FishBase to compile a table of
minimum temperature, maximum temperature, minimum depth, and maximum
depth for each species [@fishbase2024]. If a temperature range was not
explicitly listed, the preferred minimum temperature and preferred
maximum temperature (estimated from a model) were recorded
[@froese2020].

# II. Load Libraries

First, load the necessary packages for this analysis:

```{r}
#| warning: false
#| message: false
#| code-fold: true
#| results: false

#this clears out the environment
rm(list = ls())

#load necessary data packages
library(tidyverse)
library(here)
library(janitor)
library(stringr)
library(terra)
library(sf)
library(calecopal)
library(tmap)
library(ggplot2)
library(knitr)
library(tibble)
library(kableExtra)
library(htmltools)

```


# III. Define Functions {#define_functions}

All functions used in this workflow are outlined and defined in this
[section](#define_functions), and grouped into three subcategories:

-   [a) Visualization](#visualization), 
-   [b) Data Processing](#data_processing), and
-   [c) Analysis](#analysis).

## a) Visualization {#visualization}

### i) Standard Kable Format {#standard_kable}

This function can be used to generate a Kable from a tibble or dataframe that is formatted to match the general formatting in this document [@kableExtra].

The input is:

-   data: a tibble or dataframe that contains the values to be displayed in the table

-   col_names: a named vector with the names to be used for the column headers

-   caption_text: the title to be used for the table

The output is:

-   kable_output: a kable generated from the data input with the col_names as headers

    -   this kable is saved with the name "data + _kable"

    -   this output is also saved to the tables subfolder in the figures folder

```{r}
#| warning: false
#| message: false
#| code-fold: true

#write a function to generate a standard kable format 
standard_kable <- function(data, col_names, caption_text) {
  
  #extract the name of the data to save as the kable filename
  filename <- deparse(substitute(data))
  
  #write the standard formatting for the kable outputs
  kable_output <- data %>%
    kable("html",
          col.names = col_names,
          caption = htmltools::tags$div(style = "text-align: center; font-size: 20px;",
                                        htmltools::tags$strong(caption_text))) %>%
    kable_styling(full_width = FALSE, font_size = 14) %>%
    row_spec(row = 0, bold = TRUE) %>%
    kable_classic(html_font = "Times New Roman")
  
  #define the file path 
  kable_file <- file.path(tables_folder, paste0(filename, "_kable.html"))
  
  #save the kable
  save_kable(kable_output, file = kable_file)

  #print the kable
  return(kable_output)
}

```


### ii) Generate Maps {#generate_maps}

This function can be used to generate a map tmap from a SpatRaster
(only compatible with terra rasters) [@tmap; @terra]. In this workflow, this function is
used to generate a map with the "optimal" and "not optimal" land
classifications for each EEZ region for each potential finfish species.

Colors:

1.   Optimal Regions:
    - We will visualize the optimal regions in <span style="background-color:#CD4F39; padding:2px;">light red</span>. 
2.   Not Optimal Regions:
    - We will visualize the not optimal regions in <span style="background-color:#D3D3D3; padding:2px;">light grey</span>. 

The input is:

-   rast_obj: a SpatRaster (1 layer)
-   map_title: a title to be used for the map

The output is:

-   map: a tmap of the rast_obj generated for the region_name

    -   the static tmap is saved with the name "map_title + _map.png" 

    -   the interactive tmap is saved with the name "map_title + _map.html" 

    -   these outputs are also saved to the maps subfolder in the figures folder
    
```{r}
#| warning: false
#| message: false
#| code-fold: true

#assign a variable name for the optimal color
optimal_color <- "tomato3"

#assign a variable name for the not optimal color
not_optimal_color <- "lightgrey"

#write a function to generate a map for the species raster object
generate_maps <- function(rast_list, eez_data, map_title = NULL) {
  
  #create an empty list to hold the interactive and static maps
  map_list <- list()
  
  #extract the first raster from the list to be used as the base raster
  base_raster <- rast_list[[1]]
    
  #save the raster layer name as base_name
  base_name <- names(rast_list)[1]
  
  #if there is a map_title entry, use that for the title, if not, use the base_name
  title <- if (!is.null(map_title)) map_title else base_name
  
  #create a base map to be used from the first raster in the list
  combined_map <- tm_shape(base_raster) +
    tm_raster(palette = c(not_optimal_color, optimal_color),
              alpha = 0.5,
              labels = c("Not Optimal", "Optimal"),
              title = "Suitability") +
    tm_shape(eez_data) + 
    tm_fill(col = "color",
            popup.vars = c("EEZ Region" = "region"),
            alpha = 0.5,
            title = "EEZ Regions") +
    tm_layout(main.title = title,
              main.title.position = "center",
              main.title.size = 1.5, 
              main.title.fontfamily = "Times New Roman",
              main.title.fontface = "bold",
              legend.outside = TRUE)
  
  #for every object in the raster list create a map by iterating through the remaining rasters in the list
  if (length(rast_list) > 1) {
    for (i in 2:length(rast_list)) {
    
    #extract the raster from the list 
    raster_layer <- rast_list[[i]]
    
    #add the new layer to the map
    combined_map <- combined_map +
      tm_shape(raster_layer) +
      tm_raster(palette = c(not_optimal_color, optimal_color),
                alpha = 0.5,
                legend.show = FALSE)
    
    }
  }
  
  #make the map name using map_title and define file path for the static .png file
  map_name_png <- file.path(maps_folder, paste(title, "map.png", sep = "_"))
  
  #save the static tmap as PNG
  tmap_save(combined_map, filename = map_name_png)
    
  #create an interactive title
  title_int <- htmltools::tags$h1(title,
                                  style = "text-align: center; font-family: Times New Roman; font-size: 24px; font-weight: bold;")
  
  #combine the interactive title and map into an interactive map object
  map_int <- htmltools::browsable(htmltools::tagList(title_int,
                                                     tmap_leaflet(combined_map)))
  
  #make the HTML file name and define file path
  map_name_html <- file.path(maps_folder, paste0(title, "_map.html"))
  
  #save the interactive map as an HTML file
  htmltools::save_html(map_int, file = map_name_html)
  
  #add the interactive map to the map_list
  map_list$interactive_map <- map_int
  
  #add the static map to the map_list
  map_list$static_map <- map_name_png

  #return the file path for the combined interactive map and the static  file
  return(map_list)
  
}

```


### iii) Generate Heatmaps {#generate_heatmap}

This function can be used to generate a heatmap from the final region rank dataframe generated from the [rank_regions](#rank_regions) function. In this workflow, this function is used to create a heatmap which visualizes the regions that are the most suitable for aquaculture farms for both oysters and the 26 potential finfish species.

Colors:

-   We will visualize the optimal percent cover using a spectrum of colors ranging from zero percent cover in **<span style="background-color:#F0FFFF; padding:2px;">light blue</span>** to high percent cover in **<span style="background-color:#CD4F39; padding:2px;">red</span>**.

The input is:

-   rank_df: the rank dataframe generated which has a optimal percent cover and rank for the potential species 

The output is:

-   heatmap: a heatmap 

    -   the heatmap is saved with the name "filename + _heatmap.png" 

    -   this output is also saved to the heatmaps subfolder in the figures folder

```{r}
#| warning: false
#| message: false
#| code-fold: true

#assign a variable name for the optimal color
optimal_color <- "tomato3"

#write a function to generate a heatmap for the potential species and their ranking per region 
generate_heatmap <- function(rank_df) {
  
  #extract the name of the data to save as the filename
  filename <- deparse(substitute(rank_df))
  
  #create a heatmap from the rank_df
  heatmap <- ggplot(rank_df, aes(x = species, y = region, fill = percent_cover)) +
    geom_tile(color = "black", size = 0.5) +  
    geom_text(aes(label = round(percent_cover, 1)), #show the percent cover to 1 decimal
              color = "black", size = 2) +
    scale_fill_gradient2(midpoint = 0, #make the midpoint 0 so we can make the 0 percent_cover values the "azure" color
                        mid = "azure",
                        high = optimal_color,
                        name = "Percent\nCover",
                        breaks = seq(0, 5, by = 1),
                        limits = c(0, 5)) +
    scale_y_discrete(labels = c("washington" = "Washington",
                                "oregon" = "Oregon",
                                "northern_california" = "Northern California",
                                "central_california" = "Central California",
                                "southern_california" = "Southern California")) +
    labs(title = "Species vs Region Rankings",
        subtitle = "Heatmap of Percent Optimal Cover",
        x = "Species",
        y = "Region") +  
    theme(plot.title = element_text(family = "Times New Roman", 
                                    face = "bold",
                                    size = 16,
                                    hjust = 0.5), 
          plot.subtitle = element_text(family = "Times New Roman",
                                      size = 14,
                                      hjust = 0.5),
          axis.title = element_text(family = "Times New Roman",
                                    face = "bold",
                                    size = 14),
          axis.text.x = element_text(family = "Times New Roman",
                                    size = 10,
                                    angle = 45,
                                    hjust = 1,
                                    vjust = 1),
          axis.text.y = element_text(family = "Times New Roman", 
                                    size = 10),
          legend.title = element_text(family = "Times New Roman",
                                      face = "bold",
                                      size = 10),
          legend.text = element_text(family = "Times New Roman",
                                    size = 10),
          legend.position = "right",
          legend.box.background = element_rect(color = "black", size = 1))

  #define the file path 
  heatmap_file <- file.path(heatmap_folder, paste0(filename, "_heatmap.png"))
  
  #save the heatmap
  ggsave(heatmap, file = heatmap_file)
  
  #return the heatmap
  return(heatmap)

}

```


## b) Data Processing {#data_processing}

### i) Identify Spatial Properties {#spatial_properties}

This function can be used to create a tibble with all of the geometric
properties (resolution, extent, origin, and coordinate reference system)
for a list of spatial objects with the class "SpatRaster" or "sf object"
(both terra and sf compatible). This is useful for comparing the spatial
properties of the spatial objects in a workflow to ensure compatibility
before processing begins.

The input is:

-   spatial_file_list: a list of spatial objects

The output is:

-   spatial_properties: a tibble with a column for the name of the
    spatial object, the resolution, extent, origin, and coordinate
    reference system (crs)

```{r}
#| warning: false
#| message: false
#| code-fold: true

#write a function to create a tibble with all the spatial properties for a list of spatial objects
extract_spatial_properties <- function(spatial_file_list) {
  
  #create an empty tibble to hold the output
  spatial_properties <- tibble(name = character(),
                               resolution = character(),
                               extent = character(),
                               origin = character(),
                               crs = character())
  
  #extract the name of the list 
  list_name <- deparse(substitute(spatial_file_list))
  
  #write a for loop to extract the properties into the tibble for review
  for (file_name in names(spatial_file_list)) {
    
    #extract the current spatial object
    obj <- spatial_file_list[[file_name]]
    
    #since we have sf objects and SpatRasters, we must handle them differently:
  
    ##if the object is a SpatRaster
    if (inherits(obj, "SpatRaster")){
      
      #if the object is a raster stack (has more than one layer):
      if (nlyr(obj) > 1) {
        
        ##for each layer of the raster stack
        for (layer_name in names(obj)) {
          
          #extract the layer
          layer <- obj[[layer_name]] 
        
          #extract the spatial properties for this layer:
        
          ##extract the resolution (cell size) as a string with 5 decimals
          resolution <- paste(round(res(obj), 5), collapse = ", ")
        
          ##extract the extent (raster size) as a string with 2 decimals
          extent <- paste(round(ext(obj), 2), collapse = ", ")
        
          ##extract the origin (0,0) as a string with 7 decimals
          origin <- paste(round(origin(obj), 7), collapse = ", ") 
    
          ##extract the crs from the current object
          crs <- terra::crs(obj)
    
          #extract the EPSG ID part from the CRS string to make the CRS easier to read
          crs_id <- sub('.*ID\\["EPSG",(\\d+)\\].*', 'EPSG:\\1', as.character(crs))
    
          #add the new information to the tibble for raster properties
          spatial_properties <- spatial_properties %>%
            add_row(name = layer_name,
                    resolution = resolution,
                    extent = extent,
                    origin = origin,
                    crs = crs_id)
          }
        
        #if the object is a single raster:  
        } else {
          
          #extract the spatial properties for this layer:
          
          ##extract the resolution (cell size) as a string with 5 decimals
          resolution <- paste(round(res(obj), 5), collapse = ", ")
    
          ##extract the extent (raster size) as a string with 2 decimals
          extent <- paste(round(ext(obj), 2), collapse = ", ")
       
          ##extract the origin (0,0) as a string with 7 decimals
          origin <- paste(round(origin(obj), 7), collapse = ", ") 
    
          ##extract the crs from the current object
          crs <- terra::crs(obj)
    
          #extract the EPSG ID part from the CRS string to make the CRS easier to read
          crs_id <- sub('.*ID\\["EPSG",(\\d+)\\].*', 'EPSG:\\1', as.character(crs))
    
          #add the new information to the tibble for raster properties
          spatial_properties <- spatial_properties %>%
            add_row(name = file_name,
                    resolution = resolution,
                    extent = extent,
                    origin = origin,
                    crs = crs_id)
          }
      
      ##if the object is a sf object  
      } else if (inherits(obj, "sf")) {
        
        #extract the spatial properties for this layer:
    
        ##extract the EPSG ID part from the CRS string to make the CRS easier to read
        crs_id <- gsub('.*ID\\[\\\"(EPSG)\\\",(\\d+)\\].*', '\\1:\\2', as.character(crs)) 
    
        #add the new information to the tibble for raster properties
        spatial_properties <- spatial_properties %>%
          add_row(name = file_name,
                  resolution = NA,
                  extent = NA,
                  origin = NA,
                  crs = crs_id)
      }
  }
  
  #remove duplicate rows based on the "name" column that are generated for sf objects (inherently sf and dataframes)
  spatial_properties <- spatial_properties %>%
    distinct(name, .keep_all = TRUE)
  
  #create a tibble name to save
  tibble_name <- paste0(list_name, "_tibble")
  
  #rename the tibble with the list name and save to the global environment 
  return(assign(tibble_name, spatial_properties, envir = .GlobalEnv))
}

```


### ii) Check CRS {#check_crs}

This function can be used to verify that all the spatial objects in a
list (both terra and sf compatible) have a pre-defined coordinate
reference system (crs). This is necessary to ensure before any map
algebra can be performed.

The input is:

-   obj: a list of spatial objects
-   target_crs: a target crs that all the spatial objects in the list
    are compared against

The output is:

-   a logical (TRUE/FALSE) if the coordinate reference system of the
    object matches the target crs

```{r}
#| warning: false
#| message: false
#| code-fold: true

#write a function to verify that all the spatial objects have the same crs
check_crs <- function(obj, target_crs) {
  
  #if the object is a "SpatRaster"
  if (inherits(obj, "SpatRaster")) {
    
    #returns a logical (TRUE/FALSE) if the crs matches the target_crs (TRUE) or not (FALSE)
    return(terra::crs(obj) == target_crs)  
  
  #if the object is an "sf object"    
  } else if (inherits(obj, "sf")) {
    
    #returns a logical (TRUE/FALSE) if the crs matches the target_crs (TRUE) or not (FALSE)
    return(st_crs(obj)$wkt == target_crs) ##need to use wkt since terra and sf print crs differently
    
  } 
}

```


### iii) Check Spatial Properties {#check_spatial}

This function can be used to check that the geometric properties (resolution, extent, origin, and crs) of two spatial objects (compatible with terra and sf) match. This function is a modified version of the [extract_spatial_properties](#extract_spatial_properties) function.

The input is:

-   spatial_file_list: a list of spatial objects
-   ref_crs: a coordinate reference system (crs) used as the reference for comparison of list objects

The output is:

-   spatial_properties: a tibble with a column for the name of the spatial object, the resolution, extent, origin, and coordinate reference system (crs) and each entry will say "MATCH" or "DOES NOT MATCH"

```{r}
#| warning: false
#| message: false
#| code-fold: true

#write a modified version of the extract_spatial_properties function to check the spatial properties 
check_spatial_properties <- function(spatial_file_list, ref_crs) {
  
  #create an empty tibble to hold the output
  spatial_properties_check <- tibble(name = character(),
                                     resolution_match = character(),
                                     extent_match = character(),
                                     origin_match = character(),
                                     crs_match = character())
  
  #extract the name of the list 
  list_name <- deparse(substitute(spatial_file_list))
  
  #we will use the first file in the list as the reference object for verification
  ref_obj <- spatial_file_list[[1]]
  
  #now identify the spatial properties of the first object
  
  ##identify the reference resolution
  ref_resolution <- paste(res(ref_obj), collapse = ", ")

  ##identify the reference extent
  ref_extent <- paste(ext(ref_obj), collapse = ", ")
  
  ##identify the reference origin
  ref_origin <- paste(origin(ref_obj), collapse = ", ")
  
  ##identify the reference crs
  ref_crs <- ref_crs
  
  #write a for loop to extract the properties into the tibble for review
  for (file_name in names(spatial_file_list)) {
    
    #extract the current spatial object
    obj <- spatial_file_list[[file_name]]
    
    #since we have sf objects and SpatRasters, we must handle them differently:
  
    ##if the object is a SpatRaster 
    if (inherits(obj, "SpatRaster")){
      
      #if the object is a raster stack (has more than one layer):
      if (nlyr(obj) > 1) {
        
        ##for each layer of the raster stack
        for (layer_name in names(obj)) {
          
          #extract the layer
          layer <- obj[[layer_name]] 
        
          #extract the spatial properties for this layer:
        
          ##extract the resolution (cell size) as a string
          resolution <- paste(res(obj), collapse = ", ")
        
          ##extract the extent (raster size) as a string 
          extent <- paste(ext(obj), collapse = ", ")
        
          ##extract the origin (0,0) as a string
          origin <- paste(origin(obj), collapse = ", ") 
    
          ##extract the crs from the current object
          crs <- terra::crs(obj)
    
          #check if resolution, extent, origin, and CRS match the reference raster
          resolution_match <- ifelse(resolution == ref_resolution, "MATCH", "DOES NOT MATCH")
          extent_match <- ifelse(extent == ref_extent, "MATCH", "DOES NOT MATCH")
          origin_match <- ifelse(origin == ref_origin, "MATCH", "DOES NOT MATCH")
          crs_match <- ifelse(crs == ref_crs, "MATCH", "DOES NOT MATCH")
    
          #add the new information to the tibble for raster properties
          spatial_properties_check <- spatial_properties_check %>%
            add_row(name = layer_name,
                    resolution_match = resolution_match,
                    extent_match = extent_match,
                    origin_match = origin_match,
                    crs_match = crs_match)
        }
        
        #if the object is a single raster:  
        } else {
          #extract the spatial properties for this layer:
          
          ##extract the resolution (cell size) as a string
          resolution <- paste(res(obj), collapse = ", ")
        
          ##extract the extent (raster size) as a string 
          extent <- paste(ext(obj), collapse = ", ")
        
          ##extract the origin (0,0) as a string
          origin <- paste(origin(obj), collapse = ", ") 
    
          ##extract the crs from the current object
          crs <- terra::crs(obj)
    
          #check if resolution, extent, origin, and CRS match the reference raster
          resolution_match <- ifelse(resolution == ref_resolution, "MATCH", "DOES NOT MATCH")
          extent_match <- ifelse(extent == ref_extent, "MATCH", "DOES NOT MATCH")
          origin_match <- ifelse(origin == ref_origin, "MATCH", "DOES NOT MATCH")
          crs_match <- ifelse(crs == ref_crs, "MATCH", "DOES NOT MATCH")
    
          #add the new information to the tibble for raster properties
          spatial_properties_check <- spatial_properties_check %>%
            add_row(name = file_name,
                    resolution_match = resolution_match,
                    extent_match = extent_match,
                    origin_match = origin_match,
                    crs_match = crs_match)
        }
      
      ##if the object is a sf object
      } else if (inherits(obj, "sf")) {
        
        #extract the spatial properties for this layer:
    
        #extract the crs from the current object
        crs <- sf::st_crs(obj)$wkt  
        
        #check that the CRS matches the ref_crs
        crs_match <- ifelse(crs == ref_crs, "MATCH", "DOES NOT MATCH")
    
        #add the new information to the tibble for raster properties
        spatial_properties_check <- spatial_properties_check %>%
          add_row(name = file_name,
                  resolution_match = NA,
                  extent_match = NA,
                  origin_match = NA,
                  crs_match = crs_match)
      }
    }
  
  #remove duplicate rows based on the "name" column that are generated for sf objects (inherently sf and dataframes)
  spatial_properties_check <- spatial_properties_check %>%
    distinct(name, .keep_all = TRUE)
  
  #create a tibble name to save
  tibble_name <- paste0(list_name, "_tibble")
  
  #rename the tibble with the list name and save to the global environment 
  return(assign(tibble_name, spatial_properties_check, envir = .GlobalEnv))
}

```


## c) Analysis {#analysis}

### i) Identify Suitable Locations {#suitable_locations}

This function can be used to reclassify the geometries in a SpatRaster
(only compatible with terra) based on an optimal parameter range. It
generates a new SpatRaster that has three classifications:

1.  below range
2.  optimal
3.  above range

The input is:

-   raster_obj: a SpatRaster (1 layer)
-   min_parameter: a numeric value for the minimum end of the parameter
    range
-   max_paramter: a numeric value for the maximum end of the parameter
    range

The output is:

-   mask: a raster object that has been reclassified into the three
    groups and saved with the name "raster_obj + \_mask"

```{r}
#| warning: false
#| message: false
#| code-fold: true

#write a function to reclassify the geometries in a raster based on the optimal parameter range 
suitable_locations <- function(raster_obj, min_parameter, max_paramter) {
  
  #duplicate the raster that will be reclassified
  mask <- raster_obj
  
  #create a reclassification matrix to reclassify the raster based on the optimal parameter range
  rcl <- matrix(c(-Inf, min_parameter, 1, #below the minimum value ("below_range")
                  min_parameter, max_paramter, 2, #in the optimal range ("optimal")
                  max_paramter, Inf, 3), #above the maximum value ("above_range")
                ncol = 3, byrow = TRUE)
  
  #use the reclassification matrix to reclassify the duplicated raster with the rcl
  mask <- classify(mask, rcl = rcl)
  
  #assign the group to a classification (not optimal or optimal)
  levels(mask) <- data.frame(id = c(1, 2, 3),
                             cats = c("below_range", "optimal", "above_range"))
  
  #rename the mask to have the name of the raster_obj + _mask
  mask_name <- paste(names(raster_obj), "mask", sep = "_")
  names(mask) <- mask_name
  
  #return reclassified raster object
  return(mask)
  
}
  
```


### ii) Identify Optimal Locations {#optimal_locations}

This function is used to generate a new raster object from two optimized
SpatRaster (only compatible with terra). This function is formated to
work with SpatRasters that were generated from the previously defined
[suitable_locations](#suitable_locations) function. It generates a new
SpatRaster that has two classifications based on the optimal parameter
range of both SpatRasters (raster_1 and raster_2):

1.  not_optimal
2.  optimal

The input is:

-   raster_1: a SpatRaster generated with the
    [suitable_locations](#suitable_locations) function
-   raster_2: a SpatRaster generated with the
    [suitable_locations](#suitable_locations) function

The output is:

-   combined: a raster object that has been reclassified into the two
    groups based on the most conservative optimal parameter of both
    SpatRasters

```{r}
#| warning: false
#| message: false
#| code-fold: true

#write a function to generate a new raster object from two optimized rasters (this function is formatted to work with rasters generated from the suitable_locations locations defined above)
optimal_locations <- function(raster_1, raster_2){
  
  #duplicate raster_1 to serve as the combined raster
  combined <- raster_1
  
  #first assign any locations with missing data from either raster as "not optimal" group 1
  combined[raster_1 == 2 | raster_2 == 2] <- 1  
  
  #assign locations where BOTH parameters are "optimal" to group 2
  combined[raster_1 == 2 & raster_2 == 2] <- 2  
  
  #assign locations that are not identified as optimal in both rasters as "not optimal" to group 1
  combined[!(combined[] == 2)]  <- 1 
  
  #assign the categories a name (optimal or not optimal)
  levels(combined) <- data.frame(id = c(1, 2),
                                 cats = c("not_optimal", "optimal"))
  
  #return reclassified raster object
  return(combined)
  
}

```


### iii) Rasterize Geometries {#rasterize_geometries}

This function is used to rasterize the individual geometries in a sf
object based on the name of a column in the dataframe. In this workflow,
this was used to rasterize the region geometries from the EEZ dataframe.

The input is:

-   df: an sf object that has an name_column with corresponding
    geometries
-   template_raster: a SpatRaster that has the geometric properties
    compatible with other SpatRasters in the workflow
-   name_column: the name of the column in the df that will correspond
    to a new raster object being generated from the corresponding
    geometry

The output is:

-   a list with 3 sublists inside:
    1.  rast_list: a list with SpatRasters that are generated for each
        of the names in the name_column of the df
    2.  rast_stack: a stack of all the SpatRasters that are generated
        and stored in the rast_list
    3.  bbox_list: a list with boundary boxes for each of the names in
        the name_column of the df

```{r}
#| warning: false
#| message: false
#| code-fold: true

#write a function to rasterize individual geometries based on the column that has the corresponding name
rasterize_geometries <- function(df, template_raster, name_column) {
  
  #create a list of names the name column
  name_list <- unique(df[[name_column]])

  #create an empty list to store the individual layers 
  rast_list <- list()
  
  #create a duplicated empty raster from the template raster to be the raster stack
  rast_stack <- rast(template_raster, nlyrs = 0)
  
  #create an empty list to hold the bboxes 
  bbox_list <- list()
  
  #write a for loop to extract the geometry for each id and save it as a raster in the list
  for(unique_name in name_list) {
    
    #filter the df to the geometry that corresponds to the current name
    geometry <- df[df[[name_column]] == unique_name, ]
    
    #create a duplicated empty raster from the template raster to be the raster layer
    rast_layer <- template_raster
    
    #rasterize the extracted geometry
    geom_rast <- rasterize(geometry, rast_layer)
    
    #rename the raster with the name
    names(geom_rast) <- unique_name
    
    #add the new raster to the list
    rast_list[[unique_name]] <- geom_rast
    
    #create a boundary box from the geometry
    bbox <- st_bbox(geometry)
    
    #rename the bbox with the unique_name + _bbox
    bbox_name <- paste(names(unique_name), "bbox", sep = "_")
    names(bbox) <- bbox_name
    
    #add the new bbox to the list
    bbox_list[[unique_name]] <- bbox
    
    #add the new raster as a layer to the raster stack
    rast_stack <- c(rast_stack, geom_rast)
    
  }
  
  #return a list with the individual layers, the raster stack, and the bbox_list
  return(list(rast_list = rast_list, 
              rast_stack = rast_stack, 
              bbox_list = bbox_list))
  
}

```


### iv) Extract Regions {#extract_regions}

This function is used to create a mask of a SpatRaster (only compatible
with terra) for each layer in a SpatRaster stack. This function is
a modified version of the [rasterize_geometries](#rasterize_geometries)
function. In this workflow, this function is used to create a masked
SpatRaster for each region in the EEZ with the input raster
(raster_obj_input) being the SpatRaster generated for each potential
species.

The input is:

-   raster_obj_input: a SpatRaster object that is to be masked
-   rast_stack_input: a SpatRaster stack that contains layers that will
    be used to mask the raster_obj_input

The output is:

-   a list with 2 sublists inside:
    1.  rast_list: a list with masked SpatRasters that are generated for
        each of the layers in the rast_stack_input
    2.  rast_stack: a stack of all the SpatRasters that are generated
        and stored in the rast_list

```{r}
#| warning: false
#| message: false
#| code-fold: true

#write a function to create a list with a raster stack, rasters, and a map for each region 
extract_regions <- function(raster_obj_input, rast_stack_input) {
  
  #create an empty raster to store the output of the for loop
  rast_list <- list()
  
  #create a duplicate of the raster object to be the masked layer
  rast_obj <- raster_obj_input
  
  #create a duplicated empty raster from the raster object to be the raster stack
  rast_stack <- rast(rast_stack_input, nlyrs = 0)
  
  for (region_name in names(rast_stack_input)) {
    
    #extract the region layer from the raster stack
    region_layer <- rast_stack_input[[region_name]] 
    
    #use the region_layer to mask the rast_obj
    region_layer_mask <- terra::mask(rast_obj, region_layer)
    
    #rename the raster with the region_name
    names(region_layer_mask) <- region_name
    
    #add the new raster to the list
    rast_list[[region_name]] <- region_layer_mask
    
    #add the new raster as a layer to the raster stack
    rast_stack <- c(rast_stack, region_layer_mask)
  }
  
  #return a list with the individual layers and the raster stack
  return(list(rast_list = rast_list, 
              rast_stack = rast_stack))
}

  
```


### v) Rank Regions {#rank_regions}

This function is used to create a tibble that ranks the regional
SpatRasters (only compatible with terra) generated for each of the
species using the previously defined functions. The regions are ranked
based on the percent of the total area that is classified as optimal for
cultivation. In this workflow, this function ranks the regions from 1
(best EEZ region for cultivation) to 5 (worst EEZ region for
cultivation) based on a calculated optimal percent_cover.

The input is:

-   rast_stack: a SpatRaster stack that contains an EEZ region in each
    layer that has the classifications of "optimal" or "not_optimal"

The output is:

-   region_rank: a tibble with a column for the scientific name of the
    species, the EEZ region, the optimal area in that region for that
    species, the total area in that region for that species, the optimal
    percent cover in that region for that species, and the rank of that
    region for that species. It includes all 26 potential finfish
    species

```{r}
#| warning: false
#| message: false
#| code-fold: true

#write a function to extract the create a tibble with each species and the rank for their optimal regions
rank_regions <- function(rast_stack){
  
  #extract the name of the data to save as the plots filename
  filename <- deparse(substitute(rast_stack))
  
  #create a tibble to hold the output of the for loop
  region_rank <- tibble(species = character(),
                        region = character(),
                        optimal_area = numeric(),
                        total_area = numeric(),
                        percent_cover = numeric(),
                        rank = numeric())
  
  #for every layer in the rast_stack
  for (layer_name in names(rast_stack)){
    
    #extract the region layer from the raster stack
    layer <- rast_stack[[layer_name]] 
  
    #calculate the area (km^2) for each cell
    cell_area <- cellSize(layer, unit = "km")
    
    #calculate the area of all the cells with the "not_optimal" (group 1) classification
    not_optimal_area_km2 <- sum(values((layer == 1) * cell_area), na.rm = TRUE)
    
     #calculate the area of all the cells with the "optimal" (group 2) classification
    optimal_area_km2 <- sum(values((layer == 2) * cell_area), na.rm = TRUE)
    
    #calculate the total area in the region
    total_area_km2 <- optimal_area_km2 + not_optimal_area_km2
    
    #calculate the percent cover of optimal area
    percent_cover <- (optimal_area_km2 /total_area_km2) *100
    
    #add the new information to the tibble 
    region_rank <- region_rank %>%
      add_row(species = NA,
              region = layer_name,
              optimal_area = optimal_area_km2,
              total_area = total_area_km2,
              percent_cover = percent_cover,
              rank = NA)
    
  }
  
  #rank the regions with the highest optimal percent cover being ranked 1
  region_rank <- region_rank %>%
    mutate(rank = ifelse(percent_cover == 0, NA, 
                         rank(-percent_cover, ties.method = "max"))) #if the percent cover is zero assign the rank NA, and if there is a tie, assign it the max rank to be conservative
  
  #create a tibble name to save
  tibble_name <- paste0(filename, "_tibble")
  
  #rename the tibble with the filename and save a table with the rankings for each region and each species to the global environment 
  return(assign(tibble_name, region_rank, envir = .GlobalEnv))
  
}

```


# IV. Set Up {#set_up}

This [section](#set_up) sets up the filepaths and files for later
analysis:

-   [a) Define File Paths](#file_paths),
-   [b) Load Spatial Data](#spatial_data), and
-   [c) Examine Geometric Properties](#examine_geometric_properties).

## a) Define File Paths {#file_paths}

### i) Input Data {#input_file_paths}

1.  Sea Surface Temperature (SST):

    -   To characterize the average sea surface temperature inside the
        region of interest, we obtained average annual sea surface
        temperature data for the years 2008 to 2012.

    -   This data was generated by the [NOAA Coral Reef Watch Daily 5km
        Satellite Sea Surface Temperature (SST)
        (v3.1)](https://coralreefwatch.noaa.gov/product/5km/index_5km_sst.php)
        (@noaa_crw_sst_2024).

2.  Bathymetry:

    -   To characterize the depth profile inside the region of interest,
        we obtained bathymetry data from the [General Bathymetric Chart
        of the Oceans
        (GEBCO)](https://www.gebco.net/data_and_products/gridded_bathymetry_data/#area)
        (@gebco2024).

3.  Economic Exclusion Zone (EEZ):

    -   To define the area of U.S. waters that could potential be
        regions of interest for a marine aquaculture operation, we
        obtained a shapefile outlining the economic exclusion zones
        (EEZ) along the west coast of the U.S.

    -   This data was developed by the [Flanders Marine
        Institute](https://www.marineregions.org/eez.php)
        (@flanders2024).

```{r}
#| warning: false
#| message: false
#| code-fold: true

#define the pathway to access the average annual sea surface temperature (sst) files 
sst_files <- list.files(here::here("data"),
                        pattern = "average_annual_sst.*\\.tif$",
                        full.names = TRUE)

#define the pathway to access the depth file
depth_file <- here::here("data", "depth.tif")

#define the pathway to access the economic exclusion zone (eez) files 
eez_file <- here::here("data", "wc_regions_clean.shp")

#define the pathway to access the csv with the potential finfish species data
species_file <- here::here("data", "potential_finfish_species.csv")

```


### ii) Output Data {#output_file_paths}

1.  maps_folder: All maps are saved to the maps subfolder within the
    figures folder
2.  tables_folder: All tables/kables are saved to the tables subfolder
    within the figures folder
3.  heatmap_folder: All density plots are saved to the
    heatmap subfolder within the figures folder

```{r}
#| warning: false
#| message: false
#| code-fold: true

#define the pathway to save to the maps subfolder in the figures folder
maps_folder <- here::here("figures", "maps")

#define the pathway to save to the tables subfolder in the figures folder
tables_folder <- here::here("figures", "tables")

#define the pathway to save to the heatmap subfolder in the figures folder
heatmap_folder <- here::here("figures", "heatmap")

```


## b) Load Spatial Data {#spatial_data}

Now that we have defined the file paths, we can read in the necessary
spatial data:

### i) Sea Surface Temperature (SST)

-   The sea surface temperature image files were read in as a SpatRaster
    stack with terra containing a layer for each year [@terra].

```{r}
#| warning: false
#| message: false
#| code-fold: true

#read in the sea surface temperature (sst) rasters 
sst_rasters <- terra::rast(sst_files)

#write a for loop to rename each layer to have the format sst_year
for (i in seq_along(names(sst_rasters))) {
  
  #extract the year for the corresponding file, by extracting the last four digits 
  year <- str_extract(basename(sst_files[i]), "\\d{4}") #extracting the last four digits
  
  #format the name of the current layer to be sst_year and save it as the new name for the layer
  names(sst_rasters)[i] <- paste("sst", year, sep = "_")
}

```


### ii)  Bathymetry

-   The bathmetry image file was read in as a SpatRaster with terra
    containing a depth layer [@terra].

```{r}
#| warning: false
#| message: false
#| code-fold: true

#read in the depth raster
depth_raster <- terra::rast(depth_file)

#rename the layer
names(depth_raster) <- "depth"

```


### iii)  Economic Exclusion Zone (EEZ)

-   The polygons for the economic exclusion zones were read in as an sf
    object with sf and the data frame was cleaned with to only
    contain the regions and the corresponding total area (km\^2) while
    remaining sticky [@sf].

```{r}
#| warning: false
#| message: false
#| code-fold: true
#| results: false

#read in the economic exclusion zone (eez) raster
eez <- sf::st_read(eez_file)

#create a vector with region_id's that correspond to the regions based on the geographic order
region_ids <- c("Washington" = 1,
                "Oregon" = 2,
                "Northern California" = 3,
                "Central California" = 4,
                "Southern California" = 5)

#create a region_id column in the eez dataset that assigns the regions a number from north (washington) to south (southern_california)
eez <- eez %>%
  mutate(region_id = region_ids[rgn])

#create a version of the dataframe with the capitalized region names still
eez_caps <- eez %>%
  clean_names() %>%
  rename(region = rgn) %>%
  select(region_id, region, area_km2, geometry)

#tidy up the data
eez <- eez %>%
  clean_names() %>% #clean up the column names
  mutate(region = str_replace_all(rgn, " ", "_") %>% 
           str_to_lower()) %>% #create a region column with the name in snake_case
  select(region, area_km2, geometry) #only keep the necessary columns

```


## c) Examine Geometric Properties {#examine_geometric_properties}

Next, we need to examine the geometric properties (resolution, extent,
origin, and coordinate reference system (crs)) for each of the spatial
objects that we are working with. We will need to transform any spatial
objects that have mismatched properties.

### i) Identify Properties

We will start by identifying the geometric properties of each of the
spatial objects with the
[extract_spatial_properties](#spatial_properties) function and save these properties in a table with the [standard_kable](#standard_kable) function:

```{r}
#| warning: false
#| message: false
#| code-fold: true

#create a summary list with all the files
spatial_file_list <- list(sst_rasters = sst_rasters,
                          depth_raster = depth_raster,
                          eez = eez)

#identify the spatial properties of all the spatial files in the list
spatial_properties <- extract_spatial_properties(spatial_file_list)

#save the column names for the kable
col_names <- c("Name" = "Name",
               "Resolution" = "Resolution",
               "Extent" = "Extent",
               "Origin" = "Origin",
               "CRS" = "CRS")

#generate the kable from the tibble output
standard_kable(spatial_properties, col_names, caption = "Geometric Properties Identification Results")

```


It appears that the coordiante reference system for the bathymetry data
(depth_raster) and the economic exclusion zone data (eez) does not match
that of the sea surface temperature data (sst_rasters).

### ii) Transform Coordinate Reference Systems

Therefore, we will transform the bathymetry data (depth_raster) and the
economic exclusion zone data (eez) to match the coordinate reference
system (crs = EPSG:9122) of sea surface temperature data (sst_rasters).

```{r}
#| warning: false
#| message: false
#| code-fold: true

#define the target CRS (EPSG:9122) to be used in the transformations
target_crs <- crs(sst_rasters)

#transform the depth_raster to the target crs
depth_raster <- terra::project(depth_raster, target_crs)

#transform the eez to the target crs
eez <- st_transform(eez, crs = target_crs)

#create a summary list with all the files
verify_crs_list <- list(sst_rasters = sst_rasters,
                          depth_raster = depth_raster,
                          eez = eez)

```


### iii) Verify CRS Match

Now verify that all of the spatial files have the same coordinate
reference system with the [check_crs](#check_crs) function before
continuing with analysis:

```{r}
#| warning: false
#| message: false
#| code-fold: true

#check that all the files have the same crs (target_crs)
all_crs_match <- sapply(verify_crs_list, check_crs, target_crs = target_crs)

#verify that all the files have the same crs 
if (all(all_crs_match)) {
  
  #if all the files have the same crs (all TRUE) then print
  print("All spatial files have the target CRS")
  
} else {
  
  #if all the files do NOT have the same crs (at least one FALSE) then print
  print("The following files do not match the target CRS:")
  print(names(spatial_file_list)[!all_crs_match])
}

```


Since all the spatial files have the same coordinate reference system,
we can move to the analysis process.

# V. Oyster Aquaculture {#oyster_aquaculture}

This [section](#oyster_aquaculture) identifies locations that would be
suitable for only bivalve farming:

-   [a) SST and Bathymetry Data](#sst_bath_data),
-   [b) Determine Suitable Oyster Cultivation Locations](#oyster_range),
    and
-   [c) Visualize Suitable Locations](#oyster_locations).

## a) SST and Bathymetry Data {#sst_bath_data}

To start, we will ensure that the sea surface temperature (SST) and the
bathymetry data are compatible to process together.

### i) Transform SST Data

First, we need to calculate the average sea surface temperature by
taking the mean across all of the rasters (from the years 2008 to 2012),
and then convert this temperature from Kelvin to Celsius.

```{r}
#| warning: false
#| message: false
#| code-fold: true

#find the mean sea surface temperature and store it as a single layer raster
sst_mean <- terra::mean(sst_rasters, na.rm = TRUE) 

#now convert the temperature to Celcius by subtracting 273.15 to convert from Kelvin
sst_mean_celsius <- sst_mean - 273.15

#rename the layer for the celcisus raster
names(sst_mean_celsius) <- "mean_sst_c"

```


### ii) Identify Geometric Properties

Now we need to verify that the rasters are compatible to be combined. We
will start by identifying the geometric properties of with the
[extract_spatial_properties](#spatial_properties) function and save these properties in a table with the [standard_kable](#standard_kable) function:

```{r}
#| warning: false
#| message: false
#| code-fold: true

#create a summary list with all the files
sst_bath_file_list <- list(sst_mean_celsius = sst_mean_celsius,
                           depth_raster = depth_raster)

#identify the spatial properties of all the files in the list
properties_sst_bath <- extract_spatial_properties(sst_bath_file_list)

#generate the kable from the tibble output
standard_kable(properties_sst_bath, col_names, caption = "Geometric Properties Identification Results")

```


It appears that the rasters have the same coordinate reference systems
(since we transformed them to match above), but differ in their
resolution, extent, and origin.

### iii) Match Resolution, Extent, and Origin

To ensure compatible geometric properties we will transform the data so
both raster have the same resolution, extent, origin, and coordinate
reference system. Since the bathymetry data has a larger extent than the
sea surface temperature, we can crop this raster to match the extent and
origin of the temperature raster.

```{r}
#| warning: false
#| message: false
#| code-fold: true

#crop the bathymetry raster to match the extent/origin of the sst raster 
depth_raster_crop <- terra::crop(depth_raster, sst_mean_celsius)

```


Since the bathymetry data has a finer resolution than the sea surface
temperature (SST) data, we will coarsen the bathymetry data by
resampling the data using the nearest neighbor approach.

```{r}
#| warning: false
#| message: false
#| code-fold: true

#resample the cropped bathymetry data using the nearest neighbor approach 
depth_resampled <- terra::resample(depth_raster_crop, sst_mean_celsius, method = "near")

```


Now we will verify that the transformations were successful and the
resulting rasters have compatible geometric properties.

### iv) Identify Geometric Properties

Now we need to verify that the rasters are compatible to be combined by
ensuring the geometric properties have been successfully transformed
with the [check_spatial_properties](#check_spatial) function and save these properties in a table with the [standard_kable](#standard_kable) function:

```{r}
#| warning: false
#| message: false
#| code-fold: true

#create a summary list with all the files
sst_bath_transform_file_list <- list(sst_mean_celsius = sst_mean_celsius,
                                     depth_resampled = depth_resampled)

#identify the spatial properties of all the files in the list
properties_sst_bath_transform <- check_spatial_properties(sst_bath_transform_file_list, ref_crs = target_crs)

#generate the kable from the tibble output
standard_kable(properties_sst_bath_transform, col_names, caption = "Geometric Properties Identification Results")

```


Since the transformations were successful, we can move forward with
identifying suitable locations for oyster cultivation along the west
coast.

## b) Determine Suitable Oyster Cultivation Locations {#oyster_range}

Before we can identify potential finfish species that will be suitable
options for marine aquaculture, we need to determine which locations are
within the scope of the temperature and depth parameters for oysters.

### i) Temperature Conditions

To start, we will classify locations as "optimal" or "not optimal" for
oyster cultivation based on the [average sea surface
temperature](#sst_bath_data) calculated above. This is done by creating
a mask of the average sea surface temperature data (mean_sst_celcius)
and assigning appropriate categorical levels using the
[suitable_locations](#suitable_locations) function:

```{r}
#| warning: false
#| message: false
#| code-fold: true

#define the minimum temperature 
min_temp <- 11

#define the maximum temperature
max_temp <- 30

#generate a masked version of the temperature raster to identify optimal and not optimal locations
sst_mask <- suitable_locations(sst_mean_celsius, min_temp, max_temp)

```


### ii) Depth Conditions

Now we classified locations as "optimal", "not optimal", or "land" for
oyster cultivation based on the [bathymetry data](#sst_bath_data)
processed above. This is done by creating a mask of the bathymetry data
(depth_resampled) and assigning appropriate categorical levels using the
[suitable_locations](#suitable_locations) function:

```{r}
#| warning: false
#| message: false
#| code-fold: true

#define the "minimum" depth (meters below sea level)
min_depth <- -70

#define the "maximum" depth (sea surface)
max_depth <- 0

#generate a masked version of the bathymetry raster to identify optimal and not optimal locations
depth_mask <- suitable_locations(depth_resampled, min_depth, max_depth)

#we will now rename the "not optimal" group 3 to be land (elevations above sea level)
levels(depth_mask) <- data.frame(id = c(1, 2, 3),
                                 cats = c("not_optimal", "optimal", "land"))

```


### iii) Combine Optimal Conditions

Once the physical condition rasters have been reclassified, we can
combine the reclassified rasters (sst_mask and depth_mask) into a single
raster that contains the optimal locations for oyster cultivation based
on both parameters using the [optimal_locations](#optimal_locations)
function:

```{r}
#| warning: false
#| message: false
#| code-fold: true

#create a raster for the optimal geographic range for oyster cultivation based on the temperature and depth parameters
oyster_range <- optimal_locations(sst_mask, depth_mask)

```


### iv) Overlay Economic Exclusion Zones

After generating a raster with the optimal location for oyster
cultivation, we can overlay the economic exclusion zones (eez) data.
Before combining this information, we must rasterize the eez data to
make it compatible with the other SpatRaster objects. We will create an
individual raster object for each region (Southern California, Central
California, Northern California, Oregon, and Washington) of the economic
exclusion zones as well as a raster stack which contains each region as
a layer using the [rasterize_geometries](#rasterize_geometries)
function:

```{r}
#| warning: false
#| message: false
#| code-fold: true

#rasterize the regions and create a raster stack with all of the regions
eez_results <- rasterize_geometries(eez, oyster_range, "region")

#save the list of region rasters 
eez_regions <- eez_results$rast_list

#save the raster stack
eez_stack <- eez_results$rast_stack

#save the list of boundary boxes for each region 
eez_bbox_list <- eez_results$bbox_list

```


## c) Visualize Suitable Locations {#oyster_locations}

Now we can visualize the suitable locations for oyster cultivation in
each region of the economic exclusion zone (**<span style="color:#E4DECE;">Washington</span>**, **<span style="color:#ECBD95;">Oregon</span>**, **<span style="color:#9BB1BB;">Northern California</span>**, **<span style="color:#346575;">Central California</span>**, **<span style="color:#0B4221;">Southern California</span>**) along the west coast of the
U.S.

### i) EEZ Region Colors 

Let's start by making a color palette to visualize each of the
regions within the west coast EEZ using these colors. 

```{r}
#| warning: false
#| message: false
#| code-fold: true

#generate a palette from the calecopal bigsur color palette
bigsur_palette <- cal_palette(name = "bigsur")

#create a vector of selected hex codes from the bigsur_palette
region_palette <- c("#E4DECE", "#ECBD95", "#9BB1BB", "#346575", "#0B4221")

#assign the colors in the palette to region names
region_names <- c("Washington",
                  "Oregon",
                  "Northern California",
                  "Central California",
                  "Southern California")

#assign the colors in the palette to the list of names in region_names
region_palette_named <- setNames(region_palette, region_names)

#create a copy of the eez dataframe with a color that corresponds to each region in the region_palette_named
eez_color <- eez_caps %>%
  mutate(color = region_palette_named[region])

```


### ii) Generate Maps

And then we can use the [generate_maps](#generate_maps) function to create an interactive map to visualize the optimal locations (**<span style="color:#CD4F39;">optimal</span>** and **<span style="color:#D3D3D3;">not optimal</span>**)
for oyster cultivation within each EEZ region: 

```{r}
#| warning: false
#| message: false
#| code-fold: true

#use the generate maps function to create an interactive map of the suitable oyster cultivation locations 
oyster_map_results <- generate_maps(oyster_range,
                                    eez_data = eez_color,
                                    map_title = "Suitable Oyster Cultivation Locations and EEZ Regions")

#extract the interactive map from the oyster_map_results
interactive_map <- oyster_map_results$interactive_map

#print the interactive map 
interactive_map

```


# VI. Potential Finfish Aquaculture {#finfish_aquaculture}

This [section](#finfish_aquaculture) identifies locations that would be
suitable for **BOTH** bivalve and finfish aquaculture:

-   [a) Load Species Data](#load_species) and
-   [b) Identify Potential Compatible Species](#potential_species).

## a) Load Species Data {#load_species}

Start by loading in the csv with the metadata for the 26 potential
finfish species identified in the [background section](#potential_finfish_species):

```{r}
#| warning: false
#| message: false
#| code-fold: true

#read in the csv for the potential species 
potential_species <- read_csv(species_file)

#subset the data to only the species common name and scientific name
potential_species_tbl <- potential_species %>%
  select(1:2) %>%
  mutate(number = paste0(row_number())) %>% #add a number based on the row number in the df
  select(3,1,2)

#select only the first 13 rows of the table 
potential_species_tbl_1 <- potential_species_tbl %>%
  slice(1:13) 

#select only the last 13 rows of the table
potential_species_tbl_2 <- potential_species_tbl %>%
  slice(14:26)

#combine the two tables
potential_species_comb <- cbind(potential_species_tbl_1, potential_species_tbl_2)

#tidy up the data
potential_species <- potential_species %>%
  clean_names() %>%
  mutate(common_name = str_replace_all(common_name, " ", "_") %>% str_to_lower(),
         scientific_name = str_replace_all(scientific_name, " ", "_") %>% str_to_lower())

```


Then print out a table with a list of the potential species being
considered and save this output in a table with the [standard_kable](#standard_kable) function:

```{r}
#| warning: false
#| message: false
#| code-fold: true

#save the column names for the kable
col_names_p <- c("Number" = "Number",
                 "Common Name" = "Common Name",
                 "Latin Name" = "Latin Name",
                 "Number" = "Number",
                 "Common Name" = "Common Name",
                 "Latin Name" = "Latin Name")

#generate the kable from the tibble output
standard_kable(potential_species_comb, col_names_p, caption = "Potential Finfish Species") %>%
  column_spec(column = c(3, 6), italic = TRUE) #reformat for the resized table output 

```


## b) Identify Potential Compatible Species {#potential_species}

Now we will use the functions defined above to evaluate the
compatibility of each potential finfish species in each EEZ region. For
each species, we will generate a SpatRaster for each EEZ region whose
area has been classified as "optimal" or "not optimal" for the
cultivation of that species and oysters. We will then stack these
SpatRasters to generate a SpatRaster stack for each species containing a
layer for each region. Additionally we will generate a tibble which
ranks each EEZ region for each potential finfish species based on the
percent of optimal area in the region. Finally, we will also create a
map for each region layer generated for each potential species. The
functions are used in the below order:

1.  [suitable_locations](#suitable_locations),

-   this function is used twice in the workflow;
    a)  the first time is used to find the optimal location for the
        current species based on the temperature range, and
    b)  the second time is used to find the optimal location for the
        current species based on the depth range.

2.  [optimal_locations](#optimal_locations),

-   this function is used twice in the workflow;
    a)  the first time is used to find the optimal location for the
        current species based on the depth raster and temperature raster
        generated, and
    b)  the second time is used to find the optimal location for the
        current species AND oysters.

3.  [extract_regions](#extract_regions), 

-   this function is used to create a raster stack with each layer in the stack being a region in the economic exclusion zone for each species 

4.  [rank_regions](#rank_regions), and

-    this function is used to rank each region from 1 (best EEZ region for cultivation) to 5 (worst EEZ region for cultivation) based on a calculated optimal percent cover for each species and oysters. 

5.  [generate_maps](#generate_maps). 

-   this function is used to create a static and interactive map for each species and oysters in each EEZ region with the "optimal" and "not optimal" classifications

Output: This workflow will generate three outputs:

1.  species_rast_list: a list with 26 sublists (one for each species).
    Each species sublist contains two more sublists: a) rast_list: a
    list with region specific SpatRasters that are generated for each of
    the EEZ Regions b) rast_stack: a stack of all the SpatRasters that
    are generated and stored in the rast_list
2.  region_rank: a tibble with a column for the scientific name of the
    species, the EEZ region, the optimal area in that region for that
    species, the total area in that region for that species, the optimal
    percent cover in that region for that species, and the rank of that
    region for that species. It includes all 26 potential finfish
    species
3.  species_map: a list with 26 sublists (one for each species). Each
    species sublist contains a map for each EEZ region with the
    "optimal" and "not optimal" classifications

```{r}
#| warning: false
#| message: false
#| code-fold: true

#create an empty list to hold the output of the for loop
species_rast_list <- list()

#create a dataframe column with the scientific names from the potential_species_tbl
map_name_df <- potential_species_tbl[, 3]

#convert the map_name_df to a list
map_name_list <- as.list(map_name_df)

#bind the map_name_list on to the potential_species dataframe
potential_species_names <- cbind(potential_species, map_name_df)

#rename the new column to be tidyverse friendly
potential_species_names <- potential_species_names %>%
  rename(species_name_caps = 9)

#create a tibble to hold the output of the for loop
region_rank <- tibble(species = character(),
                      region = character(),
                      optimal_area = numeric(),
                      total_area = numeric(),
                      percent_cover = numeric(),
                      rank = numeric())

#create a list to hold the species maps
species_map <- list()

#for every species in the potential species dataframe
for(i in 1:nrow(potential_species_names)) {
  
  #extract the properly formatted species name from the map_name_list
  species_name_caps <- potential_species_names$species_name_caps[i]
  
  #save the species name as the scientific name
  species_name <- potential_species_names$scientific_name[i]
  
  #extract the minimum temperature for this species
  min_temp <- potential_species_names$min_temp[i]
  
  #extract the maximum temperature for this species
  max_temp <- potential_species_names$max_temp[i]
  
  #since the thermal range of tolerance was predicted from a model for some species, we have classified this in other columns (preferred_min_temp and preferred_max_temp) and left the min_temp and max_temp column empty
  
  ##so if the min_temp is NA we will use the preferred_min_temp instead
  min_temp <- ifelse(is.na(min_temp), #if the min_temp is NA
                     potential_species_names$preferred_min_temp[i], #use the preferred_min_temp
                     min_temp) #if is.na(min_temp) = FALSE then use the min_temp already extracted
  
  ##or if the max_temp is NA we will use the preferred_max_temp instead
  max_temp <- ifelse(is.na(max_temp), #if the max_temp is NA
                     potential_species_names$preferred_max_temp[i], #use the preferred_max_temp
                     max_temp) #if is.na(max_temp) = FALSE then use the max_temp already extracted
  
  #the min_depth and max_depth in the dataframe are recorded as negative numbers to represent meters below sea level to align with the format in the bathymetry data which has a vertical datum of mean sea level
  
  ##however the function wants the min_parameter value which would correspond to the max_depth, so extract the maximum depth for this species
  min_depth <- potential_species_names$max_depth[i]
  
  ##the function wants the max_parameter value which would correspond to the min_depth, so extract the minimum depth for this species
  max_depth <- potential_species_names$min_depth[i]
  
  #FUNCTION ONE A: generate a masked version of the sea surface temperature data using the suitable_locations function
  temp_rast <- suitable_locations(sst_mean_celsius, min_temp, max_temp)
  
  #FUNCTION ONE B:generate a masked version of the bathymetry using the suitable_locations function
  depth_rast <- suitable_locations(depth_resampled, min_depth, max_depth)
  
  #FUNCTION TWO A: determine the optimal range for this species using the optimal_locations function
  optimal_rast <- optimal_locations(temp_rast, depth_rast)
  
  #FUNCTION TWO B: determine the optimal range for this species AND the oysters using the optimal_locations function
  potential_locations <- optimal_locations(oyster_range, optimal_rast)

  #FUNCTION THREE: generate a raster stack with each layer in the stack being a region in the economic exclusion zone
  results_list <- extract_regions(potential_locations, eez_stack)
  
  #extract the species raster stack from the results list
  species_stack <- results_list$rast_stack
  
  #extract the species raster list from the results list
  species_list <- results_list$rast_list

  #FUNCTION FOUR: rank the regions for each species 
  species_rank <- rank_regions(species_stack)
  
  #add the species name to the tibble
  species_rank <- species_rank %>%
    mutate(species = species_name)
  
  #add the new species_rank tibble to the full tibble 
  region_rank <- bind_rows(region_rank, species_rank)
  
  #FUNCTION FIVE: generate a static and an interactive map
  map_results <- generate_maps(rast_list = species_list, 
                               eez_data = eez_color, 
                               map_title = species_name_caps)
  
  #add the map to the species_map list
  species_map[[species_name]] <- map_results$interactive_map
  
  #save the species results list to the species rast list
  species_rast_list[[species_name]] <- results_list
 
}

```


Now let's visualize some of the results from our analysis.

# VII. Results {#results}

This [section](#results) summarizes the results of the above analysis
through a variety of visualizations:

-   [a) Potential Species Region Ranking](#species_ranks),
-   [b) Ideal Locations](#ideal_locations), and
-   [c) Example Finfish Map](#example_map).

## a) Potential Species Region Ranking {#species_ranks} 

First, we can create a heatmap with the [generate_heatmap](#generate_heatmap) function to determine which regions are suitable most for aquaculture farms for both oysters and the 26 potential
finfish species. Here we can examine the optimal percent cover calculated for each species in each region with locations with **ZERO** optimal area in **<span style="background-color:#F0FFFF; padding:2px;">light blue</span>** and locations with **HIGH** percent cover of optimal area shown in a gradient of **<span style="background-color:#CD4F39; padding:2px;">red</span>**. 

```{r}
#| warning: false
#| message: false
#| code-fold: true

#make sure the regions are in order from south to north 
region_rank$region <- factor(region_rank$region, 
                             levels = c("southern_california",
                                        "central_california",
                                        "northern_california",
                                        "oregon",
                                        "washington"))


#create a heatmap of the potential species and their ranking per region 
generate_heatmap(region_rank)

```


## b) Ideal Locations {#ideal_locations}

Then we can determine which regions will be the most optimal for
creating farms for oyster and a finfish species and save these properties in a table with the [standard_kable](#standard_kable) function:

```{r}
#| warning: false
#| message: false
#| code-fold: true

#subset the region_rank dataframe to only include the regions that are the best for cultivation
top_regions <- region_rank %>%
  group_by(species) %>%
  filter(rank == 1) %>%
  ungroup()

#create a list of column names to be used in the kable
col_names_r <- c("Species" = "Species",
                 "Region" = "Region",
                 "Optimal Area" = "Optimal Area",
                 "Total Area" = "Total Area",
                 "Percent Cover" = "Percent Cover",
                 "Rank" = "Rank")

#generate the kable from the tibble output
standard_kable(top_regions, col_names_r, caption = "Top Ranked Regions for Each Species")

```


## c) Example Finfish Map {#example_map}

Finally, we can examine one of the regional maps generated for a
potential finfish species. We will visualize the maps generated for the
*Sebastes pinniger* (Canary Rockfish) as an example of the maps created
for all of the potential finfish species. We will color the regions similarly with the optimal locations (**<span style="color:#CD4F39;">optimal</span>** and **<span style="color:#D3D3D3;">not optimal</span>**)
used to visualize areas for cultivation within each EEZ region (**<span style="color:#E4DECE;">Washington</span>**, **<span style="color:#ECBD95;">Oregon</span>**, **<span style="color:#9BB1BB;">Northern California</span>**, **<span style="color:#346575;">Central California</span>**, **<span style="color:#0B4221;">Southern California</span>**): 

```{r}
#| warning: false
#| message: false
#| code-fold: true

#examine the maps generated for the sebastes_pinniger species 
tmap_mode("view")

#extract the species map from the map list
sebastes_pinniger_map <- species_map[["sebastes_pinniger"]]

#print the map
sebastes_pinniger_map

```


These results show that for all of the potential finfish species chosen,
the Washington EEZ is the most suitable for paired oyster aquaculture.
However, further parameters for each species should be examined more
closely for potential areas with the Washington region before moving
forward with development.

# IX. Citations

::: {#refs}
:::
